### 2016-07-15 stat-and-prob-letters/snb-paper.tex
Addressing comments and fixed a typo.

\documentclass[review]{elsarticle}

\usepackage{lineno}

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amsthm}
%\usepackage{url}
\usepackage{subfig}
\DeclareMathOperator*{\argmin}{argmin}
\newcommand*{\argminl}{\argmin\limits}
\newtheorem{prop}{Proposition}
\newtheorem{corollary}{Corollary}

\modulolinenumbers[5]

\journal{Statistics and Probability Letters}

%%%%%%%%%%%%%%%%%%%%%%%
%% Elsevier bibliography styles
%%%%%%%%%%%%%%%%%%%%%%%
%% To change the style, put a % in front of the second line of the current style and
%% remove the % from the second line of the style you would like to use.
%%%%%%%%%%%%%%%%%%%%%%%

%% Numbered
%\bibliographystyle{model1-num-names}

%% Numbered without titles
%\bibliographystyle{model1a-num-names}

%% Harvard
%\bibliographystyle{model2-names.bst}\biboptions{authoryear}

%% Vancouver numbered
%\usepackage{numcompress}\bibliographystyle{model3-num-names}

%% Vancouver name/year
%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}

%% APA style
%\bibliographystyle{model5-names}\biboptions{authoryear}

%% AMA style
%\usepackage{numcompress}\bibliographystyle{model6-num-names}

%% `Elsevier LaTeX' style
\bibliographystyle{elsarticle-num}
%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\begin{frontmatter}

\title{A Stopped Negative Binomial Distribution}
\tnotetext[mytitlenote]{This research was supported by grants R01CA131301, 
R01CA157749, R01CA148996, R01CA168733, and PC50CA196530 awarded by the 
National Cancer Institute along with support from the Yale Comprehensive Cancer 
Center and the Yale Center for Outcomes Research. We would also like 
to thank Rick Landin at Ignyta Inc. for his suggestions.}

%% Group authors per affiliation:
%\author{Elsevier\fnref{myfootnote}}
%\address{Radarweg 29, Amsterdam}
%\fntext[myfootnote]{Since 1880.}

%% or include affiliations in footnotes:
\author{Michelle DeVeaux}
\ead{michelle.deveaux@yale.edu}

\author{Michael J. Kane$^*$\footnote{* Corresponding author}}
\ead{michael.kane@yale.edu}

\author{Daniel Zelterman}
\ead{daniel.zelterman@yale.edu}

\address{Department of Biostatistics\\ School of Epidemiology and Public Health\\ Yale University, New Haven, CT}

\begin{abstract}
We introduce a discrete distribution suggested by curtailed
sampling rules common in early-stage clinical trials. We derive the
distribution of the smallest number of independent and identically
distributed Bernoulli trials needed to observe either $s$ successes 
or $t$ failures. This report provides a closed-form expression for the 
mass function and illustrates limiting approximations.
\end{abstract}

\begin{keyword}
discrete distribution\sep curtailed sampling
\end{keyword}

\end{frontmatter}

\linenumbers

\section{Introduction and Motivation}

Consider a prototypical early phase, single-arm clinical trial in which 
17 patients
are enrolled and treated. The binomial probability a patient 
responds to treatment is $p=0.2$ 
under the null hypothesis that the treatment is not effective.
If seven or more patients out of these 17 respond to the treatment then we 
reject this hypothesis and the treatment is deemed successful at 
significance level of $0.1$.  If fewer than seven respond then the null is not 
rejected and the treatment is deemed ineffective.

If all 17 patients are enrolled at once, as in the classic
design, then the sample size is 17. However, in most clinical trials the
patients are enrolled sequentially with one patient's outcome realized
before the next one enters the trial. In the present example, observing seven
successful patients is one endpoint and the number of enrollees required
could be as small as seven. Similarly 11
observed treatment failures also ends the trial. This sampling mechanism, in
which the experiment ends as soon as any predefined endpoint is reached, is
called {\em curtailed sampling}. Under curtailed sampling the range of the
sample size for this trial is between seven and 17.

Let us assume each patient outcome can be modeled as an independent,
identically distributed Bernoulli($p$) random variable. The trial is realized
as a sequence of Bernoulli samples stopping when either a
specified number of responders or non-responders has been reached.

\begin{figure}[bp!]
\includegraphics[width=\textwidth]{KanePlot.pdf}
\caption{
A hypothetical realization of the trial.
}
\label{fig:kane_viz}
\end{figure}

A hypothetical sample path is illustrated in Fig.~\ref{fig:kane_viz}.
The vertical axis denotes the number of
successful outcomes. The horizontal axis counts the number of patients 
enrolled. The horizontal and vertical boundaries represent
endpoints for the trial. In this case a seventh response was reached on
the $15^{\text{th}}$ enrollment.
Since the success boundary is reached, we say the treatment succeeds.

%\begin{figure}[bp!]
%\centering
%\includegraphics[width=\textwidth]{snb-first-plot.pdf}
%\caption{
%The probability mass function for a trial where $p$ 
%is 0.045 and where the trial stops
%after two patients respond or 11 patients do not respond.
%}
%\label{fig:first-plot}
%\end{figure}

\begin{figure}[bp!]
\centering
\subfloat[enrollment distribution]{\includegraphics[width=0.5\textwidth]{snb-first-plot.pdf}}
\hfill
\subfloat[mean and variance varying $p$]{\includegraphics[width=0.5\textwidth]{mean-and-variance.pdf}}
\caption{
The distribution of a trial that stops after seven patients respond to treatment
or 11 patients do not. Panel (a) shows the distribution when
$p = 0.2$. The mean and variance of the distribution varying $p$ 
between zero and one are shown in Panel (b).
}
\label{fig:exp-and-var}
\end{figure}

More generally the distribution of the number of trial enrollments is shown in 
Fig.~\ref{fig:exp-and-var} (a). There is relatively little probability mass
for values of seven through 10 since $p$ is small and it is unlikely 
the treatment will succeed quickly.
%The probability mass is concentrated at the 
%$11^{\text{th}}$ and $12^{\text{th}}$ step since $p$ 
%is small and for most realizations reach 11 
%non-responders before reaching 2 responders.
Fig.~\ref{fig:exp-and-var} (b) shows the expected value and variance for the
number of trial enrollments varying $p$ between zero and one. When $p$ is
small the probability of an individual response is small and the treatment
is more likely to fail shortly after the $11^{\text{th}}$ enrollment.
When $p$ is large the treatment is more likely to succeed and the number of 
enrollees approaches seven from above.

Since $p=1$ and $p=0$ represent the deterministic
case the corresponding variances are zero. The variance is
determined by the size of the support of the two endpoints and $p$, which 
weights the contribution of an endpoint based on how likely it is reached.
When $p$ is small it is more likely the trial will reach the failure endpoint
which has a support size of seven. As $p$ increases it becomes more likely the
trial reaches the success endpoint with a support size of 11.
The ``bump'' around $p=0.25$ results from inequality in the size of the
support of the two endpoints.

\begin{table}[t!]
\caption{Characteristics of the Stopped Negative Binomial Distribution}
\label{tab:snb}
\begin{center}
\begin{tabular}{|c|l|} \hline
Parameters & $p$ the success probability ($0\leq p \leq 1$; $q = 1-p$) \\
           & $s$ the number of successes before stopping ($s=1, 2, ...$)\\
           & $t$ the number of failures before stopping ($t=1, 2, ...$)\\ \hline
Support & min$(s,t) \leq k \leq s+t-1$  \\ \hline
$\mathbb{P}[Y=k]$ & ${k-1 \choose s-1} p^s (1-p)^{k-s} I_{\{s \leq k \leq s+t-1\}\}}+ {k-1 \choose t-1} (1-p)^s p^{k-t} I_{\{t \leq k \leq s+t-1\}}$\\ \hline
$\mathbb{P}[Y \leq k]$ & $2 - \mathcal{I}_{q}(k+1, s) - \mathcal{I}_{p}(k+1, t)$\\ 
    & where $\mathcal{I}$ is the regularized incomplete beta function.\\ \hline
Mean & $\frac{s}{p} \mathcal{I}_p(s,t) + \frac{p^{s-1} q^{t-1}}{B(s,t)} +
  \frac{t}{q} \mathcal{I}_q(t,s) + \frac{p^{t-1} q^{s-1}}{B(s,t)}$\\ 
  & where $B$ is the beta function \\ \hline
MGF & $\left(\frac{p e^x}{1 - qe^x}\right)^s 
  \mathcal{I}_{1-qe^x} (s, t) + \left(\frac{qe^x}{1-pe^x}\right)^t 
  \mathcal{I}_{1-pe^x}(t, s) $\\ \hline
Predictive Distribution & 
${k-1 \choose s-1} \frac{B\left(\alpha+s, k-s+\beta \right)}{B(\alpha, \beta)}+
{k-1 \choose t-1} \frac{B\left(\alpha + k-t, t+\beta\right)}{B(\alpha,\beta)}$\\
& for $p$ distributed as Beta($\alpha$, $\beta$). \\ \hline
\end{tabular}
\end{center}
\end{table}

In the rest of this work we derive the distribution of the number of 
enrollees needed
to observe either $s$ successes or $t$ failures. We refer to this distribution
as the Stopped Negative Binomial (SNB). Some of its characteristics are
summarized in Tab.~\ref{tab:snb}.
This paper derives these results
and explores properties of the distribution.
%The next section introduces our notation and basic results
%including the density of the distribution along with a description of
%its relation to other distributions. 
Section 2 derives the distribution function
based on a defined Bernoulli process and gives some basic properties.
Section 3 shows how the distribution is related to other standard
distributions and connects the SNB tail probability to the binomial tail 
probability.
Section 4 derives the moment generating function.
Section 5 derives the predictive distribution when $p$ has a beta distribution.

\section{Probability Mass Function}
\label{notation.section}

Let $\,b_1, b_2, \ldots \,$ denote a sequence of independent, identically
distributed, Bernoulli random variables with $\mathbb{P}[b_i=1]=p$ and
$\mathbb{P}[b_i = 0] = 1-p$, for
probability parameter $0\leq p \leq 1$. In the clinical trial setting
$\,b_i = 1$ corresponds to a patient responding to treatment.  
Let $s$ and $t$ be positive integers.  Define the SNB random
variable $Y$ as the smallest
integer value such that $\,\{b_1, \ldots , b_Y\}\,$ contains {\em either}
$\,s\,$ responders {\em or} $\,t\,$ non-responders. That is, the SNB 
distribution of $Y$ is the smallest integer such that either
$\sum_i^Y b_i = s$ or $\sum_i^Y 1-b_i = t$.

The distribution of $\,Y\,$ has support on integer values in the range
\begin{equation*}               
     \min(s,t) \leq \; Y \;\leq s+t-1  \label{range.y.eq}.
\end{equation*}
The probability mass function is
\begin{equation} \label{eqn:pmf}
\mathbb{P} [Y=k] = S(k, p, s) \ I_{\{s \leq k \leq s+t-1\}} + 
  S(k, 1-p, t) \ I_{\{ t \leq k \leq s+t-1 \}}
\end{equation}
where $I_{\{f\}}$ is the {\em indicator function}, taking the value 
of one if $f$ is true and zero otherwise, and
\begin{equation} \label{eqn:N}
S(k, p, s) = {k-1 \choose s-1} p^s (1-p)^{k-s} 
\end{equation}
is the negative binomial probability mass.

To prove (\ref{eqn:pmf}), consider the
process $\mathbf{X} = \left\{X(k) : k = 0,1,... \right\}$
with $X(0)=0$ and
\begin{equation*} \label{eqn:proc}
X_{k+1} = X_k + b_{k+1} \ I_{\{ k-t < X_k < s\}}.
\end{equation*}
At each step a patient's outcome is measured. In Fig.~\ref{fig:kane_viz} 
we consider a graphical illustration of the plot $X_k$ against
$k$. If the outcome of the $k$th patient responds to treatment then the process 
advances diagonally in the positive horizontal and vertical direction. 
If the $k$th patient does not respond
then the sample path advances in the positive horizontal direction only. The
process continues until either $X_k = s$ or $X_k = k-t$.

\begin{prop}
The distribution of the stopping time
\begin{equation*}
Y = \argminl_k \left[X_k \geq s \cup X_k \leq k-t \right]
\end{equation*}
is given at (\ref{eqn:pmf}).
\end{prop}
\begin{proof}
%The proof will proceed in two parts. First, a combinatorial justification 
%will be given for the probability mass value on each element of the support. 
%Second, it will be shown that the sum of the masses over the support sums to 
%one.

The probability a given realization of $\mathbf{X}$ reaches $s$ at
the $k$th outcome is the probability that, at time $k-1$, there are $s-1$
successful outcomes and $k-s$ unsuccessful outcomes multiplied by
the probability of a final success at time $k$. This expression is given
in (\ref{eqn:N}). 
Similarly, the probability a given realization reaches $k-t$
is the probability that, at outcome $k-1$, there are $k-t$ successful outcomes
and $t-1$ unsuccessful outcomes multiplied by the probability of a final
unsuccessful outcome at time $k$.

%Next, define
%\begin{equation} \label{stop_t}
%S'(k, p, t) = {k-1 \choose k-t} p^{k-t} (1-p)^t
%\end{equation}
%and notice that $S(k, p, s) = S'(k, 1-p, s)$ by writing
%\begin{equation*}
%{k-1 \choose k-s} = {k-1 \choose s-1}.
%\end{equation*}

To show that (\ref{eqn:pmf}) sums to one, define
%\begin{align} \label{eqn:sum_proof}
%R &= \sum_{k=s}^{s+t-1} S(k, p, s) + \sum_{k=t}^{s+t-1} S(k, 1-p, t) \\
%  &= \sum_{k=s}^{s+t-1} {k-1 \choose s-1} p^s (1-p)^{k-s} + \sum_{k=t}^{s+t-1} {k-1 \choose k-t} p^{k-t} (1-p)^t
%\end{align}
\begin{equation*} 
R = \sum_{k=s}^{s+t-1} S(k, p, s) + \sum_{k=t}^{s+t-1} S(k, 1-p, t).
\end{equation*}
If we substitute $i=k-s$ in the first summation and $j=k-t$ in the second then
$R$ can be written as the cumulative distribution function of two
negative binomial distributions:
\begin{equation} \label{eqn:transformed_sum}
R = \sum_{i=0}^{t-1} {i+s-1 \choose i} p^s (1-p)^i \; + \;
\sum_{j=0}^{s-1} {j+t-1 \choose j} p^j (1-p)^t.
\end{equation}

Let $\mathcal{I}_p(s, t)$ be the {\em regularized incomplete beta function} 
\citep{Olver2010}. This function satisfies 
$\mathcal{I}_p(s, t) = 1-\mathcal{I}_{1-p}(t, s)$ \citep{Uppuluri1970}. Then
\begin{align*}
R = \sum_{i=0}^{t-1} &{i+s-1 \choose i} p^s (1-p)^i +
\sum_{j=0}^{s-1}  {j+t-1 \choose j} p^j  (1-p)^t \\
   &= 1-\mathcal{I}_p(s, t) + 1 - \mathcal{I}_{1-p}(t, s) \\
   &= 1. 
\end{align*}
This completes the proof that (\ref{eqn:pmf}) is the distribution of the
stopping time and is a valid probability mass function.
\end{proof}

Next, we consider an interim analysis of a clinical trial after $s'$ 
patients respond to treatment 
and $t'$ do not for $s' < s$ and $t' < t$.
\begin{corollary} \label{conditional_distribution}
The number of subsequent enrollments needed 
to reach either endpoint behaves as SNB($p$, $s-s'$, $t-t'$).
\end{corollary}
Having observed $s'$ responders and $t'$ non-responders, there are $s-s'$ 
more responders needed to reach the success endpoint and $t-t'$ more 
non-responders needed to reach the failure endpoint.

\section{Connections and Approximations to Other Distributions}

\begin{figure}[p!]
\begin{center}
\includegraphics[width=\textwidth]{shapes.pdf}
\end{center}
\caption{Different shapes of the SNB distribution with parameters 
($p$, $s$, $t$), as given. Black indicates mass contributed by reaching
$s$ responders before $t$ non-responders. Grey indicates
mass contributed by reaching $t$ non-responders first. \label{shapes.fig}}
\end{figure}

The SNB is a generalization of the negative 
binomial distribution. When $t$ is large $Y-s$ has a 
negative binomial distribution with
\begin{equation*}                                    %   (1)   
\mathbb{P}[Y=s+j\ |\ t \text{ is large\ }]        \label{nb1.eq}          
  = {{s+j-1}\choose{s-1}} p^s (1-p)^j
\end{equation*}
for $\,j=0, 1,\ldots\,$. A similar statement can be made when $s$ is large
and $t$ is small. As a result, with proper parameter choice, the SNB
can mimic other probability distributions in a manner similar to 
those described in \cite{Peizer1968} and \cite{Best1974}. Examples are
shown in in Fig.~\ref{shapes.fig}.

The SNB is the 
number of enrollment outcomes to observe {\em either} $s$ responders {\em or} 
$t$ non-responders.
The SNB generalizes both the minimum (riff-shuffle) and maximum negative
binomial distributions up to a translation of the support.
For the special case of $\,s=t,$ the distribution of $\,Y\,$ is the
riff-shuffle, or minimum negative binomial distribution~\citep{Uppuluri1970}.
The maximum negative binomial \cite{Zhang2000,Zelterman2005} is the smallest 
number of outcomes necessary to 
observe at least $s$ responders {\em and} $s$ non-responders and is equivalent
to a translated version of the riff-shuffle.

%\section{Connection Between the SNB and the Binomial Tail Probability}

%There is a close connection between the tail probabilities of the SNB and the 
%binomial distributions.
%The probability of reaching the success endpoint in an 
%SNB($p$, $s$, $t$) random variable is 
%equal to the probability of at least $s$ successes in a binomial distributed 
%random variable with size $s+t-1$ and success probability $p$. 
%Likewise, the probability of reaching the failure endpoint is equal
%to the probability of at most $s-1$ successes in a binomial distribution. 
There is an equivalence between the probability of reaching an endpoint 
in the SNB model and the tail probability of the binomial distribution.
That is, the probability that the number of responders is at least $s$ in the 
binomial model is the same as the probability the treatment succeeds 
(reaches $s$) in the SNB 
model.
\begin{prop} \label{binomial_tail}
Let $Y$ be distributed as SNB($p$, $s$, $t$) and recall that $X_Y$ corresponds
to the number of responders in a trial. Let 
$B$ be distributed binomial with size $n=s+t-1$ and response probability
$p$. Then
\begin{equation}
\mathbb{P}[B \geq s] = \mathbb{P} [X_Y = s].
\end{equation}
\end{prop}
\begin{proof}
The binomial tail probability is
\begin{equation*}
\mathbb{P}[B \geq s] = 1 - \mathcal{I}_{1-p}.
\end{equation*}
%\begin{align*}
%\mathbb{P}[B \geq s] &= \sum_{k=s}^{s+t-1} {n \choose k} p^k (1-p)^{n-k} \\
%  &= 1 - \sum_{k=0}^{s-1} {n \choose k} p^k (1-p)^{n-k} \\
%  &= 1 - \mathcal{I}_{1-p}.
%\end{align*}
The corresponding success probability is
\begin{equation*}
\mathbb{P} [X_Y = s] 
  = \sum_{k=s}^{s+t-1} {k-1 \choose s-1} p^s (1-p)^{k-s}.
\end{equation*}
Let $i=k-s$. Since
\begin{equation*}
{i+s-1 \choose s-1} = {i+s-1 \choose i},
\end{equation*}
the last summation can be rewritten as
\begin{align}
\mathbb{P} [X_Y = s] &= \sum_{i=0}^{t-1} {i+s-1 \choose i} p^s (1-p)^i\\
  &= 1 - \mathcal{I}_{1-p}(t, s)
\end{align}
completing the proof.
\end{proof}

To illustrate this result, let us return to our initial example
where $s=7$, $t=11$, and $p=0.2$.  The probability masses in
Fig.~\ref{fig:snb_bin_compare} represented in 
black are equal in panels (a) and (b) as are the masses in grey.
The probability that $s$
responders are reached in the SNB process is the same as the binomial 
probability of at least seven responders. Likewise, the probability that $t$ 
non-responders are reached in the SNB process is the same as the binomial
probability of zero through six responders.

\begin{figure}[t!]
\centering
\subfloat[SNB distribution]{\includegraphics[width=0.5\textwidth]{snb_density.pdf}}
\hfill
\subfloat[binomial distribution]{\includegraphics[width=0.5\textwidth]{bin_density.pdf}}
\caption{
SNB(0.2, 7, 11) with mass contributed from 
7 responders (black) or 11 non-responders (grey) along with 
Binomial(0.2, 17) with at least 2 responders (black) or fewer (grey).
}
\label{fig:snb_bin_compare}
\end{figure}

\section{The Moment Generating Function}

The moment generating function for the SNB is calculated in a manner similar to 
that of two negative binomial distributions. 
\begin{prop} Let $Y$ be distributed SNB with parameters $p$, $s$, and $t$.
Then the moment generating function (MGF) of $Y$ is
\begin{equation} \label{eqn:mgf}
\mathbb{E}~e^{xY} = \left(\frac{p e^x}{1 - qe^x}\right)^s 
  \mathcal{I}_{1-qe^x} (s, t) + \left(\frac{qe^x}{1-pe^x}\right)^t 
  \mathcal{I}_{1-pe^x}(t, s)
\end{equation}
for $q = 1-p$ defined for $x \leq \min \left\{\log(1/p), \log(1/q) \right\}$.
\end{prop}
\begin{proof}
The MGF of the SNB is:
\begin{equation*}
\mathbb{E}~e^{xY} = \sum_{k=s}^{s+t-1} {k-1 \choose s-1} p^s q^{k-s} e^{kx} 
  + \sum_{k=t}^{s+t-1} {k-1 \choose t-1} p^{k-t} q^t e^{kx}
\end{equation*}
and can be rewritten as:
\begin{equation} \label{eqn:first_sum}
\mathbb{E}~e^{xY} = \sum_{k=s}^{s+t-1}{k-1 \choose s-1} (pe^x)^{s} (qe^x)^{k-s} 
  + \sum_{k=t}^{s+t-1}{k-1 \choose t-1} (qe^x)^t (pe^x)^{k-t}.
\end{equation}
The first summation in (\ref{eqn:first_sum}) satisfies
\begin{align*}
\sum_{k=s}^{s+t-1}{k-1 \choose s-1} (pe)^{sx} (qe^x)^{k-s} &= 
  \left(\frac{pe^x}{1 - qe^x}\right)^s \ \ \sum_{k=s}^{s+t-1} {k-1 \choose s-1} 
    (qe^x)^{k-s} (1-qe^x)^s \\
  &= \left(\frac{pe^x}{1 - qe^x}\right)^s \mathcal{I}_{1-qe^x}(s, t).
\end{align*}
Since the incomplete beta function's subscript parameter has support on zero 
to one, we have $qe^x \leq 1$. This also shows we must restrict
$x \leq -\log(q)$.
A similar expression can be derived for the second summation in 
(\ref{eqn:first_sum}) and results in
the constraint $x \leq -\log(p)$.
\end{proof}

The SNB's ability to approximate the geometric, normal, gamma, and poisson
distributions follow from it generalizing the negative binomial distribution. 
To recover the MGF of the negative binomial consider the case where
$t \rightarrow \infty$ in (\ref{eqn:mgf}). The regularized incomplete
beta function in the first term goes to one and zero in the second term.
We are left with the MGF of the negative binomial distribution.

When $s=1$ (and $t$ is still large) the SNB's MGF is the same as that of the
geometric distribution. The negative
binomial can therefore be seen as a sum of i.i.d. geometric distributions.
For an appropriately large number of samples the central limit theorem
yields a normal approximation.

Drawing connections to the gamma and poisson distributions are more complicated.
However a connection to the gamma distribution well-studied problem in the 
literature (see \cite{Ord1968,Guenther1972,Best1974} for examples). 
A connection to the poisson appears in \cite{Anscombe1950} where
it is shown that if the mean of a poisson is proportional
to a $\chi^2_{2k}$ distribution then the negative binomial is obtained. 
Both of these approximations work by equating cumulants and then showing 
that differences between betwen the cumulant generating
functions converge to zero.

The lower-half normal distribution can be approximated by setting $s=t$
for appropriately large $s$ and $t$ and $p = 0.5$. In this
case the SNB can be viewed as identical, negative binomials
approximating a normal and truncated at the median.

\section{The Predictive Probability Distribution}

Let us consider the SNB distribution where $p$ has a beta distribution.
\begin{prop} \label{prop:bayesian}
The predictive distribution of the SNB
distribution when $p$ is distributed Beta($\alpha$, $\beta$) is
\begin{align} \label{eqn:posterior}
\mathbb{P} [Y = k | s, t, \alpha, \beta ] &= 
  {k-1 \choose s-1} \frac{B\left(\alpha+s, k-s+\beta \right)}{B(\alpha, \beta)} 
    \ I_{\{s \leq k \leq s+k-1\}} \ + \nonumber \\
  & {k-1 \choose t-1} 
    \frac{B\left(\alpha + k - t, t+\beta\right)}{B(\alpha, \beta)} 
    \ I_{\{t \leq k \leq s+k-1\}}.
\end{align}
\end{prop}
\begin{proof}
For notational simplicity, assume that $k$ is in the range
$\min(s,t) \leq k \leq s+t-1$. When 
this is not the case appropriate terms should be removed as dictated by the indicator functions.
\begin{align*}
f(k | s, t, \alpha, \beta) = \frac{1}{B(\alpha, \beta)} & \int_0^1 {k-1 \choose s-1} p^{\alpha +s -1} \left(1-p\right)^{k-s+\beta-1} + \\
 & {k-1 \choose t-1} p^{k-t+\alpha-1}\left(1-p\right)^{t+\beta-1} dp \\
= \frac{1}{B(\alpha, \beta)}  {k-1 \choose s-1} & \int_0^1  p^{\alpha +s -1} \left(1-p\right)^{k-s+\beta-1} dp + \\
 & \frac{1}{B(\alpha, \beta)} {k-1 \choose t-1} \int_0^1  p^{k-t+\alpha-1}\left(1-p\right)^{t+\beta-1} dp.
\end{align*}
The result in (\ref{eqn:posterior}) follows by the definition of the 
beta function.
\end{proof}

To better understand the predictive distribution, let us once again return to 
our example where $s=7$, $t=11$. However, we assume $p$ is a beta distribution
with shape parameters $\alpha = 0.2 c$ and 
$\beta = 0.8 c$ for any $c > 0$.
This parameterization allows us to examine the relationship between 
uncertainty in $p$ and the SNB.

\begin{figure}[h!]
\centering
\subfloat[mean]{\includegraphics[width=0.5\textwidth]{bayesian-sample-expectation.pdf}}
\hfill
\subfloat[variance]{\includegraphics[width=0.5\textwidth]{bayesian-sample-variance.pdf}}
\caption{
The mean and variance of the predictive SNB with $s=7$, $t=11$
with $\mathbb{E}\ p = 0.2$ varying the uncertainty 
in the distribution of $p$. A larger value of $c$
corresponds to a beta with smaller variance.
}
\label{fig:bayesian-sample-size}
\end{figure}

Fig.~\ref{fig:bayesian-sample-size} shows how the mean
and variance of the predictive distribution is related to the SNB's 
mean and variance. As $c$ increases the 
predictive distribution sample statistics converge to corresponding 
statistics for the SNB with $p = 0.2$. 
%It may be noted that the compound 
%distribution's mean
%approaches the deterministic limit slowly. This
%is because when $c$ is small, the distribution of $p$ is heavily skewed. 
%As $c$ increases this effect is minimized.
When the distribution of $p$ is centered close to zero or one the mean 
estimate approach the deterministic limit relatively slowly in the uncertainty
parameter $c$.  Because of this, we note that for estimation it may be 
preferable to select shape parameters based on the median or mode of the 
beta distribution since they are less sensitive to distributional uncertainty.

%\section{Discussion}
%
%The SNB can be used to extend clinical trial methodology
%in at least two areas. First, it can be used to construct nested criteria
%for early stopping and efficacy. Current approaches are often inspired by
%the Simon 2-stage Phase II clinical trial \cite{Simon1989} where a
%a small number of patients are enrolled for an initial stage. For patients,
%success or failure is assessed. If at least a pre-specified number of
%successes are observed then the trial enters a second stage with more
%patients. If a sufficient number of successes are not observed then the
%trial fails.

%The SNB is being used to generalize the Simon design by providing separate
%criteria for futility and efficacy.
%The Simon design tests for a single response probability and time-to-event
%parameter for both the first
%and second stage. The SNB is being used to generalize this design by 
%allowing these parameters to vary for the two stages, and even allow
%information about patients in the first stage to be used in the second
%stage. Early results show the design provides
%both early testing for futility along with the ability
%to make statistically sound decisions sooner than traditional trials.

%A second avenue for methodological innovation comes directly from
%Corollary \ref{conditional_distribution} and Proposition \ref{prop:bayesian}.
%Our knowledge of the response probability $p$ comes from published reports
%yielding a prior distribution.
%While a trial is being conducted the response probabilities are 
%updated as new results are received. The number of individual successes
%and failure are used to learn the shape characteristics of the corresponding
%beta distribution and resulting posterior
%distribution. An estimate of the probability of reaching the success (and
%failure) endpoints can be provided at each step of the trial.

%There are also several avenues for extending the SNB itself. First is the 
%development of a stopped negative multinomial distribution allowing
%for more than two endpoints. This extension would have clinical trial
%applications including dose escalation where the outcomes are response
%to a therapy, failure to respond, and an adverse outcome endpoint where
%an enrollee is removed from the trial with neither response or non-response
%due to side effects.

%A second, more theoretical extension is to the asymptotic setting. Under
%proper reparameterization, the process from which the SNB is derived 
%can be shown to converge to a Brownian motion. In this case, we conjecture
%that the SNB can be defined in terms of the L\'{e}vy-Bachelier 
%\citep{Bachelier1900} formula
%thereby relating the SNB to the classic result in stochastic calculus.

\section*{References}

\bibliography{mybibfile}

\end{document}